---
title: "Gladiator_phyloseq"
author: "Ping Hu"
date: "2022-11-15"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(vegan)
library(ggplot2)
library(phyloseq)
library(DESeq2)
library(ComplexHeatmap)
library(ANCOMBC)
library(ggpubr)
```

## read count file into phyloseq object

```{r file, echo=FALSE}
filename="gladiator.metaphlan3.count10.7.bacteria"
test <- read.table(filename, header = TRUE, row.names = 1, sep="\t")
test[is.na(test)]<-0
d=dim(test)
X=apply(test, 1, sum)
otu_count=test[(X/d[2] >=10),]
dim(otu_count)
meta<-data.frame(colnames(otu_count))
colnames(meta)="sampleID"
meta2 <- meta %>%
  select(sampleID) %>% 
  separate(sampleID, into=c("Study", "Treat", "Visit", "SID", "asfs", "age", "id2", "Visit2", "Flowcell", "GSSid"),sep="[.]")

meta2$TreatVisit=paste0(meta2$Treat, meta2$Visit)
rownames(meta2)=colnames(otu_count)
meta2$asfs =as.numeric(meta2$asfs)
meta2$age =as.numeric(meta2$age)


taxonomy <- data.frame(row.names(otu_count), row.names(otu_count)) 
colnames(taxonomy)=c("OTU", "Taxon")
tax <- taxonomy %>%
  select(Taxon) %>% 
  separate(Taxon, into=c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"),sep="[|]")
tax.clean <- data.frame(row.names = row.names(otu_count),
                        Kingdom = str_replace(tax[,1], "k__",""),
                        Phylum = str_replace(tax[,2], "p__",""),
                        Class = str_replace(tax[,3], "c__",""),
                        Order = str_replace(tax[,4], "o__",""),
                        Family = str_replace(tax[,5], "f__",""),
                        Genus = str_replace(tax[,6], "g__",""),
                        Species = str_replace(tax[,7], "s__",""),
                        stringsAsFactors = FALSE)

tax.clean[is.na(tax.clean)] <- ""
tax.clean[tax.clean=="__"] <- ""

TAX = tax_table(as.matrix(tax.clean))
OTU = otu_table(as.matrix(otu_count), taxa_are_rows = TRUE)
SAMPLE <- sample_data(meta2)
ps <- phyloseq(OTU, TAX, SAMPLE)

```

```{r 1observed, echo=FALSE}
set.seed(111) # keep result reproductive
ps.rarefied = rarefy_even_depth(ps, rngseed=1, sample.size=1103, replace=F)
png(filename=paste0(filename,".richness.TreatVisit.png"),  width=600, height=380)
plot_richness(ps.rarefied, x="TreatVisit", measures=c("Observed", "Shannon")) +
  geom_boxplot() +
  theme_classic() +
  theme(strip.background = element_blank(), axis.text.x.bottom = element_text(angle = -90))
dev.off()
############################################################
rich = estimate_richness(ps.rarefied, measures = c("Observed", "Shannon"))
wilcox.observed <- pairwise.wilcox.test(rich$Observed, 
                                        sample_data(ps.rarefied)$TreatVisit, 
                                        p.adjust.method = "BH")
tab.observed <- wilcox.observed$p.value %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "group1") %>%
  gather(key="group2", value="p.adj", -group1) %>%
  na.omit()
print("observed richness wilcox paired test")
tab.observed
```

## shannon diversity

```{r 2shannon, echo=FALSE}
wilcox.shannon <- pairwise.wilcox.test(rich$Shannon, 
                                       sample_data(ps.rarefied)$TreatVisit, 
                                       p.adjust.method = "BH")
tab.shannon <- wilcox.shannon$p.value %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "group1") %>%
  gather(key="group2", value="p.adj", -group1) %>%
  na.omit()
print("shannon richness wilcox paired test")
tab.shannon
```

##PcoA Plot and adonis test

```{r 3PCoA bray, echo=FALSE}
dist = phyloseq::distance(ps.rarefied, method="bray")
ordination = ordinate(ps.rarefied, method="PCoA", distance=dist)
png(filename=paste0(filename,".PcoA.bray.Visit.facetTreat.png"),  width=600, height=380)
plot_ordination(ps.rarefied, ordination, color="Visit") + 
  theme_bw()+
  #theme_classic() +
  theme(strip.background = element_blank())+
  stat_chull(aes(color=Visit, fill=Visit), alpha=0.1, geom="polygon")+
  facet_grid(. ~ Treat)
dev.off()
print("adonis test for TreatVisit")
metadata <- data.frame(sample_data(ps.rarefied))
test.adonis <- adonis(dist ~ TreatVisit, data = metadata)
test.adonis <- as.data.frame(test.adonis$aov.tab)
test.adonis
```

##PAIRWISE PERMANOVA treatvisit with bray curtis distance

```{r 4pairwise permanova, echo=FALSE}
print("PAIRWISE PERMANOVA treatvisit with bray curtis distance")
cbn <- combn(x=unique(metadata$TreatVisit), m = 2)
p <- c()

for(i in 1:ncol(cbn)){
  ps.subs <- subset_samples(ps.rarefied, TreatVisit %in% cbn[,i])
  metadata_sub <- data.frame(sample_data(ps.subs))
  permanova_pairwise <- adonis(phyloseq::distance(ps.subs, method = "bray") ~ TreatVisit, 
                               data = metadata_sub)
  p <- c(p, permanova_pairwise$aov.tab$`Pr(>F)`[1])
}

p.adj <- p.adjust(p, method = "BH")
p.table <- cbind.data.frame(t(cbn), p=p, p.adj=p.adj)
p.table
```

## jaccard analysis

```{r 5PCoA jaccard, echo=FALSE}
dist = phyloseq::distance(ps.rarefied, method="jaccard")
ordination = ordinate(ps.rarefied, method="PCoA", distance=dist)

png(filename=paste0(filename,".PcoA.jaccard.Visit.facetTreat.png"),  width=600, height=380)
plot_ordination(ps.rarefied, ordination, color="Visit") + 
  theme_bw()+
  theme(strip.background = element_blank())+
  stat_chull(aes(color=Visit, fill=Visit), alpha=0.1, geom="polygon")+
  facet_grid(. ~ Treat)
dev.off()
print("PAIRWISE PERMANOVA treatvisit with jaccard distance")
p <- c()
for(i in 1:ncol(cbn)){
  ps.subs <- subset_samples(ps.rarefied, TreatVisit %in% cbn[,i])
  metadata_sub <- data.frame(sample_data(ps.subs))
  permanova_pairwise <- adonis(phyloseq::distance(ps.subs, method = "jaccard", binanry=TRUE) ~ TreatVisit, 
                               data = metadata_sub)
  p <- c(p, permanova_pairwise$aov.tab$`Pr(>F)`[1])
}

p.adj <- p.adjust(p, method = "BH")
p.table <- cbind.data.frame(t(cbn), p=p, p.adj=p.adj)
p.table
```

## NMDS and ANOSIM

```{r 6jaccard NMDS, echo=FALSE}
dist = phyloseq::distance(ps.rarefied, method="jaccard", binary = TRUE)
ordination = ordinate(ps.rarefied, method="NMDS", distance=dist)

png(filename=paste0(filename,".NMDS.jaccard.Visit.facetTreat.png"),  width=600, height=380)
plot_ordination(ps.rarefied, ordination, color="Visit") + 
  theme_bw() +
  theme(strip.background = element_blank())+
  stat_chull(aes(color=Visit, fill=Visit), alpha=0.1, geom="polygon")+
  facet_grid(. ~ Treat)
dev.off()
###################################################
print("ANOSIM jaccard")
metadata <- data.frame(sample_data(ps.rarefied))
anosim(dist, metadata$TreatVisit)
```

##PAIRWISE ANOSIM with jacarrd distance 

```{r 7PCoA anosim, echo=FALSE}
print("PAIRWISE ANOSIM with jacarrd distance")
cbn <- combn(x=unique(metadata$TreatVisit), m = 2)
p <- c()

for(i in 1:ncol(cbn)){
  ps.subs <- subset_samples(ps.rarefied, TreatVisit %in% cbn[,i])
  metadata_sub <- data.frame(sample_data(ps.subs))
  permanova_pairwise <- anosim(phyloseq::distance(ps.subs, method="jaccard", binary = TRUE), 
                               metadata_sub$TreatVisit)
  p <- c(p, permanova_pairwise$signif[1])
}

p.adj <- p.adjust(p, method = "BH")
p.table <- cbind.data.frame(t(cbn), p=p, p.adj=p.adj)
p.table
```

## barplot

```{r 8barplot, echo=FALSE}
ps.rel = transform_sample_counts(ps, function(x) x/sum(x)*100)
matrix=as.matrix(data.frame(otu_table(ps.rel)))
write.csv(matrix, file=paste0(filename, ".Species.relativeAbundance.csv"))
# agglomerate taxa
glom <- tax_glom(ps.rel, taxrank = 'Phylum', NArm = FALSE)
ps.melt <- psmelt(glom)
# change to character for easy-adjusted level
ps.melt$Phylum <- as.character(ps.melt$Phylum)

ps.melt <- ps.melt %>%
  group_by(TreatVisit, Phylum) %>%
  mutate(average=mean(Abundance))
# select group median > 1
keep <- unique(ps.melt$Phylum[ps.melt$average > 1])
ps.melt$Phylum[!(ps.melt$Phylum %in% keep)] <- "< 1"
#to get the same rows together
ps.melt_sum <- ps.melt %>%
  group_by(Sample,TreatVisit,Phylum) %>%
  summarise(Abundance=sum(Abundance))

png(filename=paste0(filename,".Phylum.png"),  width=1800, height=380)
ggplot(ps.melt_sum, aes(x = Sample, y = Abundance, fill = Phylum)) + 
  geom_bar(stat = "identity", aes(fill=Phylum)) + 
  labs(x="", y="%") +
  facet_wrap(~TreatVisit, scales= "free_x", nrow=1) +
  theme_classic() + 
  theme(strip.background = element_blank(), 
        #axis.text.x.bottom = elment_text(angle = -90),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()
        )
dev.off()
############################################################
##genus level
ps.rel = transform_sample_counts(ps, function(x) x/sum(x)*100)
# agglomerate taxa
glom <- tax_glom(ps.rel, taxrank = 'Genus', NArm = FALSE)
ps.melt <- psmelt(glom)
# change to character for easy-adjusted level
ps.melt$Genus <- as.character(ps.melt$Genus)

ps.melt <- ps.melt %>%
  group_by(TreatVisit, Genus) %>%
  mutate(average=mean(Abundance))
# select group mean > 1
keep <- unique(ps.melt$Genus[ps.melt$average > 1])
ps.melt$Genus[!(ps.melt$Genus %in% keep)] <- "< 1%"
#to get the same rows together
ps.melt_sum <- ps.melt %>%
  group_by(Sample,TreatVisit,Genus) %>%
  summarise(Abundance=sum(Abundance))
png(filename=paste0(filename,".Genus.png"),  width=1800, height=380)
ggplot(ps.melt_sum, aes(x = Sample, y = Abundance, fill = Genus)) + 
  geom_bar(stat = "identity", aes(fill=Genus)) + 
  labs(x="", y="%") +
  facet_wrap(~TreatVisit, scales= "free_x", nrow=1) +
  theme_classic() + 
  theme(legend.position = "right", 
        strip.background = element_blank(), 
        #axis.text.x.bottom = element_text(angle = -90),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
dev.off()
##########################################
############################################################
##species level
ps.rel = transform_sample_counts(ps, function(x) x/sum(x)*100)
# agglomerate taxa
glom <- tax_glom(ps.rel, taxrank = 'Species', NArm = FALSE)
ps.melt <- psmelt(glom)
# change to character for easy-adjusted level
ps.melt$Species <- as.character(ps.melt$Species)

ps.melt <- ps.melt %>%
  group_by(TreatVisit, Species) %>%
  mutate(average=mean(Abundance))
# select group mean > 1
keep <- unique(ps.melt$Species[ps.melt$average > 1])
ps.melt$Species[!(ps.melt$Species %in% keep)] <- "< 1%"
#to get the same rows together
ps.melt_sum <- ps.melt %>%
  group_by(Sample,TreatVisit,Species) %>%
  summarise(Abundance=sum(Abundance))
png(filename=paste0(filename,".Species.png"),  width=1800, height=380)
ggplot(ps.melt_sum, aes(x = Sample, y = Abundance, fill = Species)) + 
  geom_bar(stat = "identity", aes(fill=Species)) + 
  labs(x="", y="%") +
  facet_wrap(~TreatVisit, scales= "free_x", nrow=1) +
  theme_classic() + 
  theme(legend.position = "right", 
        strip.background = element_blank(), 
        #axis.text.x.bottom = element_text(angle = -90),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
dev.off()
```
##abundance analysis - deseq2

```{r 9deseq2 analysis, echo=FALSE}
sample_data(ps)$TreatVisit <- as.factor(sample_data(ps)$TreatVisit) # factorize for DESeq2
ps.taxa <- tax_glom(ps, taxrank = 'Species', NArm = FALSE)
# #################################################
# pairwise comparison between ABL and AW3
ps.taxa.sub <- subset_samples(ps.taxa, TreatVisit %in% c("ABL", "AW3"))
# filter sparse features, with > 90% zeros
ps.taxa.pse.sub <- prune_taxa(rowSums(otu_table(ps.taxa.sub) == 0) < ncol(otu_table(ps.taxa.sub)) * 0.9, ps.taxa.sub)
ps_ds = phyloseq_to_deseq2(ps.taxa.pse.sub, ~ TreatVisit)
# use alternative estimator on a condition of "every gene contains a sample with a zero"
ds <- estimateSizeFactors(ps_ds, type="poscounts")
ds = DESeq(ds, test="Wald", fitType="parametric")
alpha = 0.05
resA = results(ds, alpha=alpha)
# pairwise comparison between CBL and CW3
ps.taxa.sub <- subset_samples(ps.taxa, TreatVisit %in% c("CBL", "CW3"))
# filter sparse features, with > 90% zeros
ps.taxa.pse.sub <- prune_taxa(rowSums(otu_table(ps.taxa.sub) == 0) < ncol(otu_table(ps.taxa.sub)) * 0.9, ps.taxa.sub)
ps_ds = phyloseq_to_deseq2(ps.taxa.pse.sub, ~ TreatVisit)
# use alternative estimator on a condition of "every gene contains a sample with a zero"
ds <- estimateSizeFactors(ps_ds, type="poscounts")
ds = DESeq(ds, test="Wald", fitType="parametric")
alpha = 0.05
resC = results(ds, alpha=alpha)
# pairwise comparison between BBL and BW3
ps.taxa.sub <- subset_samples(ps.taxa, TreatVisit %in% c("BBL", "BW3"))
# filter sparse features, with > 90% zeros
ps.taxa.pse.sub <- prune_taxa(rowSums(otu_table(ps.taxa.sub) == 0) < ncol(otu_table(ps.taxa.sub)) * 0.9, ps.taxa.sub)
ps_ds = phyloseq_to_deseq2(ps.taxa.pse.sub, ~ TreatVisit)
# use alternative estimator on a condition of "every gene contains a sample with a zero"
ds <- estimateSizeFactors(ps_ds, type="poscounts")
ds = DESeq(ds, test="Wald", fitType="parametric")
alpha = 0.05 
resB = results(ds, alpha=alpha)

df_deseq = data.frame(cbind(resA$baseMean,resA$log2FoldChange, resA$pvalue, resA$padj, 
                      resB$baseMean,resB$log2FoldChange, resB$pvalue,resB$padj, 
                      resC$baseMean,resC$log2FoldChange, resC$pvalue, resC$padj))
rownames(df_deseq)=rownames(resA)
colnames(df_deseq)=c("ABL_AW3.baseMean","ABL_AW3.log2FoldChange","ABL_AW3.pval", "ABL_AW3.qval",
                   "BBL_BW3.baseMean","BBL_BW3.log2FoldChange","BBL_BW3.pval", "BBL_BW3.qval",
                   "CBL_CW3.baseMean","CBL_CW3.log2FoldChange","CBL_CW3.pval", "CBL_CW3.qval")
write.csv(df_deseq, file=paste0(filename, ".deseq-wald-parametric.csv"))


```
## ANCOM-BC

```{r 10ANCOM-BC, echo=FALSE}
ps.taxa.sub <- subset_samples(ps.taxa, TreatVisit %in% c("CBL", "CW3"))
# ancombc
out <- ancombc(phyloseq = ps.taxa.sub, formula = "TreatVisit", 
               p_adj_method = "holm", zero_cut = 0.90, lib_cut = 1000, 
               group = "TreatVisit", struc_zero = TRUE, neg_lb = TRUE, tol = 1e-5, 
               max_iter = 100, conserve = TRUE, alpha = 0.05, global = TRUE)

res2 <- out$res
samp_frac = out$samp_frac
# Replace NA with 0
samp_frac[is.na(samp_frac)] = 0 
# Add pesudo-count (1) to avoid taking the log of 0
log_obs_abn = log(out$feature_table + 1)
# Adjust the log observed abundances
log_corr_abn = t(t(log_obs_abn) - samp_frac)
# Show the first 6 samples
LogCorrectAbundanceC=round(log_corr_abn, 2)

ps.taxa.sub <- subset_samples(ps.taxa, TreatVisit %in% c("ABL", "AW3"))
out <- ancombc(phyloseq = ps.taxa.sub, formula = "TreatVisit", 
               p_adj_method = "holm", zero_cut = 0.90, lib_cut = 1000, 
               group = "TreatVisit", struc_zero = TRUE, neg_lb = TRUE, tol = 1e-5, 
               max_iter = 100, conserve = TRUE, alpha = 0.05, global = TRUE)

res3 <- out$res
samp_frac = out$samp_frac
# Replace NA with 0
samp_frac[is.na(samp_frac)] = 0 
# Add pesudo-count (1) to avoid taking the log of 0
log_obs_abn = log(out$feature_table + 1)
# Adjust the log observed abundances
log_corr_abn = t(t(log_obs_abn) - samp_frac)
# Show the first 6 samples
LogCorrectAbundanceA=round(log_corr_abn, 2)

ps.taxa.sub <- subset_samples(ps.taxa, TreatVisit %in% c("BBL", "BW3"))
# ancombc
out <- ancombc(phyloseq = ps.taxa.sub, formula = "TreatVisit", 
               p_adj_method = "holm", zero_cut = 0.90, lib_cut = 1000, 
               group = "TreatVisit", struc_zero = TRUE, neg_lb = TRUE, tol = 1e-5, 
               max_iter = 100, conserve = TRUE, alpha = 0.05, global = TRUE)

res <- out$res
samp_frac = out$samp_frac
# Replace NA with 0
samp_frac[is.na(samp_frac)] = 0 
# Add pesudo-count (1) to avoid taking the log of 0
log_obs_abn = log(out$feature_table + 1)
# Adjust the log observed abundances
log_corr_abn = t(t(log_obs_abn) - samp_frac)
# Show the first 6 samples
LogCorrectAbundanceB=round(log_corr_abn, 2)
 


df_ab=data.frame(cbind(LogCorrectAbundanceA, LogCorrectAbundanceB, LogCorrectAbundanceC))
write.csv(df_ab, file=paste0(filename, ".ancombc_log_corr_abundance.csv"))
df_lfc = data.frame(cbind(res3$p_val, res3$q_val, res3$diff_abn,
                    res$p_val, res$q_val, res$diff_abn,
                    res2$p_val, res2$q_val, res2$diff_abn)) 
                    #as.factor(tax_table(ps.taxa)[, "Species"]), check.names = TRUE)
colnames(df_lfc)=c("ABL_AW3.pval", "ABL_AW3.qval", "ABL_AW3.diff_abn","BBL_BW3.pval", "BBL_BW3.qval", "BBL_BW3.diff_abn", "CBL_CW3.pval", "CBL_CW3.qval", "CBL_CW3.diff_abn")
write.csv(df_lfc, file=paste0(filename, ".ancombc.csv"))
```
##Now is the part do not use phyloseq -- traditional way with vegan 

```{r 11No_phyloseq, echo=FALSE}
test_d=data.frame(t(otu_count))
dim(test_d)
min(apply(test_d, 1, sum)) ###301
test_a <- decostand(test_d, method = "total")
write.csv(t(test_a), file =paste0(filename, "recal.Percent.csv"))

Cname=rownames(test_d)

Clen=length(Cname) ##there are 3 annotation columns
splitname<-strsplit(Cname, "[.]")


Treat=rep("NA", Clen)
sid=rep("NA", Clen)
Visit=rep("NA", Clen)
TreatVisit=rep("NA", Clen)
asfs=rep("NA", Clen)
age=rep("NA", Clen)
id2=rep("NA", Clen)
asfsGroup=rep("NA",Clen)
ageGroup=rep("NA", Clen)
for(mm in  1:Clen ){
  Treat[mm]=splitname[[mm]][2]
  Visit[mm]=splitname[[mm]][3]
  sid[mm]=splitname[[mm]][4]
  asfs[mm]=as.numeric(splitname[[mm]][5])
  age[mm]=splitname[[mm]][6]
  id2[mm]=splitname[[mm]][7]
  TreatVisit[mm]=paste0(Treat[mm], Visit[mm])
  if(age[mm]<40){ageGroup[mm]="40Less"}
  else if (age[mm]<50){ageGroup[mm]="50-40"}
  else if (age[mm]<60){ageGroup[mm]="60-50"}
  else{ageGroup[mm]="70+-60"}
  if(asfs[mm]>=24){asfsGroup[mm]="3D"}
  else if(asfs[mm]<=10){asfsGroup[mm]="1ND"}
  else{asfsGroup[mm]="2MID"}
}
shannon<-diversity(test_d)
simp<-diversity(test_d, "simpson")
invsimp<-diversity(test_d, "inv")
#unbias.simp <- rarefy(test_d, 2) - 1
observed<-apply(test_d>0,1,sum)
N <- apply(test_d,1,sum)
Menhinick_index<-observed/sqrt(N)
Margalef_index <-(observed-1)/log(N)
mydata=data.frame(shannon,simp,invsimp,  observed, Menhinick_index, Margalef_index, 
                   as.numeric(asfs), asfsGroup, TreatVisit,Treat, Visit, age, ageGroup,sid, id2 )

write.csv(t(mydata), file =paste0(filename, "alphadiversity.csv"))
```
## 

```{r 12alpha shannon, echo=FALSE}

library(ggplot2)

#https://www.datanovia.com/en/blog/ggpubr-how-to-add-p-values-generated-elsewhere-to-a-ggplot/
library(tidyverse)
library(ggpubr)
library(rstatix)

png(filename=paste0(filename,"shannon.asfsGroup.ttest.png"))
stat.test <- mydata %>%
  #group_by(Country) %>%
  t_test(shannon ~ asfsGroup) %>%
  #adjust_pvalue() %>%
  mutate(y.position = 3.8)
bxp <- ggboxplot(mydata, x = "asfsGroup", y = "shannon", color = "asfsGroup")
bxp + stat_pvalue_manual(stat.test,   label = "p", tip.length = 0.01, step.increase = 0.1)
dev.off()

png(filename=paste0(filename,"shannon.asfsGroup.wilcox.png"))
stat.test <- mydata %>%
  #group_by(Country) %>%
  wilcox_test(shannon ~ asfsGroup) %>%
  #adjust_pvalue() %>%
  mutate(y.position = 3.8)
bxp <- ggboxplot(mydata, x = "asfsGroup", y = "shannon", color = "asfsGroup")
bxp + stat_pvalue_manual(stat.test,   label = "p", tip.length = 0.01, step.increase = 0.1)
dev.off()

png(filename=paste0(filename,"shannon.TreatVisit.wilcox.png"))
stat.test <- mydata %>%
  #group_by(Country) %>%
  wilcox_test(shannon ~ TreatVisit) %>%
  #adjust_pvalue() %>%
  mutate(y.position = 3.8)
bxp <- ggboxplot(mydata, x = "TreatVisit", y = "shannon", color = "TreatVisit")
bxp + stat_pvalue_manual(stat.test,   label = "p", tip.length = 0.01, step.increase = 0.1)
dev.off()

wilcox.shannon <- pairwise.wilcox.test(mydata$shannon, 
                                        mydata$TreatVisit, 
                                        p.adjust.method = "BH")
tab.shannon <- wilcox.shannon$p.value %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "group1") %>%
  gather(key="group2", value="p.adj", -group1) %>%
  na.omit()
print("shannon -- TreatVisit wilcoxon adjustp")
tab.shannon

```
## observed 

```{r 13alpha observed, echo=FALSE}
png(filename=paste0(filename,"shannon.TreatVisit.ttest.png"))
stat.test <- mydata %>%
  #group_by(Country) %>%
  t_test(shannon ~ TreatVisit) %>%
  #adjust_pvalue() %>%
  mutate(y.position = 3.8)
bxp <- ggboxplot(mydata, x = "TreatVisit", y = "shannon", color = "TreatVisit")
bxp + stat_pvalue_manual(stat.test,   label = "p", tip.length = 0.01, step.increase = 0.1)
dev.off()

png(filename=paste0(filename,"observed.TreatVisit.ttest.png"))
stat.test <- mydata %>%
  t_test(observed ~ TreatVisit) %>%
  #adjust_pvalue() %>%
  mutate(y.position = 15.8)
bxp <- ggboxplot(mydata, x = "TreatVisit", y = "observed", color = "TreatVisit")
bxp + stat_pvalue_manual(stat.test,   label = "p", tip.length = 0.01, step.increase = 0.1)
dev.off()

png(filename=paste0(filename,"observed.TreatVisit.wilcox.png"))
stat.test <- mydata %>%
  wilcox_test(observed ~ TreatVisit) %>%
  #adjust_pvalue() %>%
  mutate(y.position = 15.8)
bxp <- ggboxplot(mydata, x = "TreatVisit", y = "observed", color = "TreatVisit")
bxp + stat_pvalue_manual(stat.test,   label = "p", tip.length = 0.01, step.increase = 0.1)
dev.off()
########add in to show statistical table directly
wilcox.observed <- pairwise.wilcox.test(mydata$observed, 
                                        mydata$TreatVisit, 
                                        p.adjust.method = "BH")
tab.observed <- wilcox.observed$p.value %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "group1") %>%
  gather(key="group2", value="p.adj", -group1) %>%
  na.omit()
print("observed -- TreatVisit wilcoxon adjustp")
tab.observed

```
## ASFS Score

```{r 14asfs, echo=FALSE}
colnames(mydata)[7]="asfs"
png(filename=paste0(filename,"asfs.TreatVisit.ttest.png"))
stat.test <- mydata %>%
  #group_by(Country) %>%
  t_test(asfs ~ TreatVisit) %>%
  #adjust_pvalue() %>%
  mutate(y.position = 55)
bxp <- ggboxplot(mydata, x = "TreatVisit", y = "asfs", color = "TreatVisit")
bxp + stat_pvalue_manual(stat.test,   label = "p", tip.length = 0.01, step.increase = 0.1)
dev.off()

png(filename=paste0(filename,"asfs.TreatVisit.wilcox.png"))
stat.test <- mydata %>%
  #group_by(Country) %>%
  wilcox_test(asfs ~ TreatVisit) %>%
  #adjust_pvalue() %>%
  mutate(y.position = 55)
bxp <- ggboxplot(mydata, x = "TreatVisit", y = "asfs", color = "TreatVisit")
bxp + stat_pvalue_manual(stat.test,   label = "p", tip.length = 0.01, step.increase = 0.1)
dev.off()
wilcox.asfs <- pairwise.wilcox.test(mydata$asfs, 
                                        mydata$TreatVisit, 
                                        p.adjust.method = "BH")
tab.asfs <- wilcox.asfs$p.value %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "group1") %>%
  gather(key="group2", value="p.adj", -group1) %>%
  na.omit()
print("asfs -- TreatVisit wilcoxon adjustp")
tab.asfs

```
## alpha diversity for Kruskal Wallis Test

```{r 15PCoA bray, echo=FALSE}
KP_asfsGroup=kruskal.test(shannon ~ asfsGroup, data = mydata)$p.value
KP_TreatVisit=kruskal.test(shannon ~TreatVisit, data = mydata)$p.value
KP_ageGroup=kruskal.test(shannon ~ ageGroup, data = mydata)$p.value
KP_Treat=kruskal.test(shannon ~Treat, data = mydata)$p.value
KP_Visit=kruskal.test(shannon ~Visit, data = mydata)$p.value
print(paste0("KP_asfsGroup=", KP_asfsGroup, "; KP_TreatVisit=", KP_TreatVisit,"; KP_Treat=", KP_Treat,"; KP_Visit=", KP_Visit, "; KP_ageGroup=", KP_ageGroup
     ))
```
## beta diversity

```{r 16PCoA bray, echo=FALSE}
x<-test_d
beta_dist <- vegdist(x, index = "bray")
mds <- metaMDS(beta_dist)
mds_data <- as.data.frame(mds$points)
mds_data$SampleID <- rownames(mds_data)
splitname<-strsplit(mds_data$SampleID, "[._]")


mds_data$asfs=rep("NA", Clen)
mds_data$asfsGroup=rep("NA", Clen)
mds_data$age=rep("NA", Clen)
mds_data$ageGroup=rep("NA", Clen)
mds_data$sid=rep("NA", Clen)
mds_data$TreatVisit=rep("NA", Clen)
mds_data$Visit=rep("NA", Clen)
mds_data$Treat=rep("NA", Clen)
mds_data$id2=rep("NA", Clen)
splitname<-strsplit(Cname, "[._]")
for(mm in  1:Clen ){
  mds_data$Treat[mm]=splitname[[mm]][2]
  mds_data$Visit[mm]=splitname[[mm]][3]
  mds_data$sid[mm]=splitname[[mm]][4]
  mds_data$asfs[mm]=as.numeric(splitname[[mm]][5])
  mds_data$age[mm]=splitname[[mm]][6]
  mds_data$id2[mm]=splitname[[mm]][7]
  mds_data$TreatVisit[mm]=paste0(mds_data$Treat[mm], mds_data$Visit[mm])
  if(mds_data$age[mm]<40){mds_data$ageGroup[mm]="40Less"}
  else if (mds_data$age[mm]<50){mds_data$ageGroup[mm]="50-40"}
  else if (mds_data$age[mm]<60){mds_data$ageGroup[mm]="60-50"}
  else{mds_data$ageGroup[mm]="70+-60"}
  if(mds_data$asfs[mm]>=24){mds_data$asfsGroup[mm]="3D"}
  else if(mds_data$asfs[mm]<=10){mds_data$asfsGroup[mm]="1ND"}
  else{mds_data$asfsGroup[mm]="2MID"}
  
}
 png(filename=paste0(filename,".beta.asfsGroup.png"),  width=600, height=380)
 p1<-ggplot(mds_data, aes(x = MDS1, y = MDS2, color = asfsGroup)) +geom_point(size=6)+theme_bw()+stat_chull(aes(color=asfsGroup, fill=asfsGroup), alpha=0.1, geom="polygon")
 print(p1)
 dev.off()
 
 png(filename=paste0(filename,".beta.asfsGroup.facetTreat.png"),  width=600, height=380)
 p1<-ggplot(mds_data, aes(x = MDS1, y = MDS2, color = asfsGroup)) +geom_point(size=3)+theme_bw()+stat_chull(aes(color=asfsGroup, fill=asfsGroup), alpha=0.1, geom="polygon")+ facet_grid(. ~ Treat)
 print(p1)
 dev.off()
 
png(filename=paste0(filename,".beta.Visit.facetTreat.png"),  width=600, height=380)
 p1<-ggplot(mds_data, aes(x = MDS1, y = MDS2, color = Visit)) +geom_point(size=6)+theme_bw()+stat_chull(aes(color=Visit, fill=Visit), alpha=0.1, geom="polygon")+ facet_grid(. ~ Treat)
 print(p1)
 dev.off()
 
 bc.dist <- vegdist(test_d, method = "bray")
jaccard.dist <- vegdist(test_d, method = "jaccard")
mds2 <- metaMDS(jaccard.dist)
mds_data2 <- as.data.frame(mds2$points)
mds_data2$SampleID <- rownames(mds_data2)
mds_data2$asfs=mds_data$asfs
mds_data2$asfsGroup=mds_data$asfsGroup
mds_data2$age=mds_data$age
mds_data2$ageGroup=mds_data$ageGroup
mds_data2$sid=mds_data$sid
mds_data2$TreatVisit=mds_data$TreatVisit
mds_data2$Visit=mds_data$Visit
mds_data2$Treat=mds_data$Treat
mds_data2$id2=mds_data$id2
png(filename=paste0(filename,".beta.Visit.facetTreat.jaccard.png"),  width=600, height=380)
p1<-ggplot(mds_data2, aes(x = MDS1, y = MDS2, color = Visit)) +geom_point(size=6)+theme_bw()+stat_chull(aes(color=Visit, fill=Visit), alpha=0.1, geom="polygon")+ facet_grid(. ~ Treat)
print(p1)
dev.off()
# cluster communities using average-linkage algorithm
bc.clust <- hclust(bc.dist, method = "average")
# plot cluster diagram
png(filename=paste0(filename,".bray-Curtis-cluster.png"), width=8000, height=2000)
plot(bc.clust, ylab = "Bray-Curtis dissimilarity")
dev.off()
print("adonis result bc.dist ~ TreatVisit")
adonis2(bc.dist ~ TreatVisit)
test.adonis <- adonis(bc.dist ~ TreatVisit)
test.adonis <- as.data.frame(test.adonis$aov.tab)
test.adonis

```
## PAIRWISE PERMANOVA

```{r 17PAIRWISE PERMANOVA, echo=FALSE}
#PAIRWISE PERMANOVA
cbn <- combn(x=unique(TreatVisit), m = 2)
p <- c()

for(i in 1:ncol(cbn)){
  ps.subs <- test_d [c(TreatVisit %in% cbn[,i]), ]
  metadata_sub <- mds_data[c(TreatVisit %in% cbn[,i]), ]
  
  permanova_pairwise <- adonis(vegdist(ps.subs, method = "bray") ~ metadata_sub$TreatVisit)
  p <- c(p, permanova_pairwise$aov.tab$`Pr(>F)`[1])
}

p.adj <- p.adjust(p, method = "BH")
p.table <- cbind.data.frame(t(cbn), p=p, p.adj=p.adj)
p.table
```
## adonis result jaccard.dist ~ TreatVisit

```{r 18, echo=FALSE}
print("adonis result jaccard.dist ~ TreatVisit")
adonis2(jaccard.dist ~ TreatVisit)
test.adonis <- adonis(jaccard.dist ~ TreatVisit)
test.adonis <- as.data.frame(test.adonis$aov.tab)
test.adonis
```
## adonis result jaccard.dist ~ TreatVisit

```{r 19, echo=FALSE}
#PAIRWISE PERMANOVA
cbn <- combn(x=unique(TreatVisit), m = 2)
p <- c()

for(i in 1:ncol(cbn)){
  ps.subs <- test_d [c(TreatVisit %in% cbn[,i]), ]
  metadata_sub <- mds_data[c(TreatVisit %in% cbn[,i]), ]
  
  permanova_pairwise <- adonis(vegdist(ps.subs, method = "jaccard") ~ metadata_sub$TreatVisit)
  p <- c(p, permanova_pairwise$aov.tab$`Pr(>F)`[1])
}

p.adj <- p.adjust(p, method = "BH")
p.table <- cbind.data.frame(t(cbn), p=p, p.adj=p.adj)
p.table

```
## anosim result jaccard.dist ~ TreatVisit

```{r 20, echo=FALSE}
print("Anosim Test of Jaccard distance to Treatvisit")
anosim(jaccard.dist, TreatVisit)
```
## pairwise anosim

```{r 21, echo=FALSE}
print("Pairwise Anosim Test")
cbn <- combn(x=unique(TreatVisit), m = 2)
p <- c()

for(i in 1:ncol(cbn)){
  ps.subs <- test_d [c(TreatVisit %in% cbn[,i]), ]
  metadata_sub <- mds_data[c(TreatVisit %in% cbn[,i]), ]
  
  permanova_pairwise <- anosim(vegdist(ps.subs, method = "jaccard") , metadata_sub$TreatVisit)
  p <- c(p, permanova_pairwise$signif[1])
}

p.adj <- p.adjust(p, method = "BH")
p.table <- cbind.data.frame(t(cbn), p=p, p.adj=p.adj)
p.table

```
